\bf Visual diagnostics for constrained optimisation with application to guided tours

### 2020-10-09 paper/paper.Rmd
write up noisy index diff seeds example

```{r load-data}
files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))

source(here::here("source/indice.r"))
source(here::here("source/sim_data.r"))
```

@@

Figure \ref{toy-search} presents a comparison between the two plotting options discussed above. Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations. The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.

```{r toy-search, fig.asp=0.5,  fig.cap="\\label{toy-search}A comparison of \`search_better\` with different parameter value on \`max_tries\`. A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting. With \`max_tries = 500\` specification, the algorithm is working much harder towards to the end."}
bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
          holes_2d_better %>% mutate(max_tries = 25)) %>% 
  explore_trace_search(group = max_tries) + 
  scale_color_botanical(palette = "daisy") +
  facet_wrap(vars(max_tries),labeller = "label_both", ncol = 2) +
  xlab("iteration")
```

@@

Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, ...

@@

@@

The second experiment with the noisy index is to understand how the optimisers perform when there are multiple informative variables in the case of 1D projection. The dataset used is `boa6` with `x2` and `x7` being informative. The two possible theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)` repectively and hence, the global maximum happens when variable `x7` is found.

A simulation with 20 different seeds shows that `search_better` can find either one of the informative variable but there's no guarantee on which seed finds which variable. `search_better_random` manages to find the global maximum via a metropolis-hasting algorithm without being affected by the seed. To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature $t_0$ so that the probability of accepting an inferior basis is not too high.  Figure \ref{pca-better-default}, .... Figure \ref{pca-better-random-default} shows the comparison, in terms of the trace plot and PCA plot, between `search_better_random` and two `search_better` algorithms where one finds `x2` and another finds `x7`.  The trace plot shows that the global maximum has been found by `search_better_random` and one `search_better`, but not the other. The PCA plot further illustrate how the two algorithms finds the global maximum: `search_better_random` travels along the top right corner on the projected 2D circle while the path of `search_better` is much shorter and only goes pass in between the two theoretical bases.

(there's more things going on about both algorithms experimenting with different seeds.)

@@

there's more text here Figure \ref{pca-better-random-tuned}, Comparison between search_better_random and two different search_better. In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases. serach_better_random and one of the serach_better have found the global maximum but the other fails. The lower panel shows the PCA plot for each of the three algorithms. The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum. The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path. This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance

@@

>>>>

```{r external, include = FALSE, cache = FALSE}
library(knitr)
files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
read_chunk(here::here('scripts/vis-diag.R'))
```

@@

Figure \@ref(fig:toy-search) presents a comparison between the two plotting options discussed above. Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations. The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.

```{r toy-search, fig.cap = "(ref:toy-search)"}
```

(ref:toy-search) A comparison of `search_better` with different parameter value on `max_tries`. A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting. With `max_tries = 500` specification, the algorithm is working much harder towards to the end.

@@

@@

Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, There should actually be more writing here to and also the this examples may need to be changed.

@@

@@

@@

The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is  presented. The dataset used is `boa6` with `x2` and `x7` being informative. The two possible theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)` repectively and hence, the global maximum happens when variable `x7` is found. Simulation is done with 20 randomly generated seeds on different parameter specifications on two optimsers: `search_better` and `search_better_random`. The data obbject is collected for each simulation and principal component analysis is applied to project the bases as points into 2D space with the orthonormality constraint imposed on the 6D space preserved by randomly generated basis on the surface of the 6D sphere.

Figure \ref{pca-better-default} shows that half of the seeds end up getting trapped in the local maximum on optimiser `search_better`. This happens because `search_better` will always accept a basis with higher index value. Figure \ref{pca-better-random-default} shows the results when the optimiser `search_better_random` is used. With a probability acceptance on inferior basis, more simulation ends up finds the global maximum. An evidence of the superior of this probabilistic implementation is simulation 5107 (row 1 column 2 in Figure \ref{pca-better-default} and row 3 column 5 in Figure \ref{pca-better-random-default}). When using `search_better` to optimise, it goes straight down the track to find `V2` whilst with `search_better_random`, it first has a tendency to hand towards `V2` but then makes a sharp turn back to the track that finds `V7`.

In the next simulation, as in Figure \ref{pca-better-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7. Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated basis and hence represents a wider search. By this specification, all the 20 seends find the global maximum and a few simulations i.e. seed 4761 (row 4,  column 5), 9209 (row 3, column 2), 2986 (row 2, column 3) from this double guarantee of wider search and probabilistic acceptance.

### 2020-10-09 paper/paper.tex
write up noisy index diff seeds example

Figure \ref{toy-search} presents a comparison between the two plotting options discussed above. Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations. The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-search-1.pdf}
\caption{\label{fig:toy-search}\label{toy-search}A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}. A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting. With \texttt{max\_tries\ =\ 500} specification, the algorithm is working much harder towards to the end.}
\end{figure}

@@

Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, \ldots{}

@@

The second experiment with the noisy index is to understand how the optimisers perform when there are multiple informative variables in the case of 1D projection. The dataset used is \texttt{boa6} with \texttt{x2} and \texttt{x7} being informative. The two possible theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235 repectively and hence, the global maximum happens when variable \texttt{x7} is found.

A simulation with 20 different seeds shows that \texttt{search\_better} can find either one of the informative variable but there's no guarantee on which seed finds which variable. \texttt{search\_better\_random} manages to find the global maximum via a metropolis-hasting algorithm without being affected by the seed. To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature \(t_0\) so that the probability of accepting an inferior basis is not too high. Figure \ref{pca-better-default}, \ldots{}. Figure \ref{pca-better-random-default} shows the comparison, in terms of the trace plot and PCA plot, between \texttt{search\_better\_random} and two \texttt{search\_better} algorithms where one finds \texttt{x2} and another finds \texttt{x7}. The trace plot shows that the global maximum has been found by \texttt{search\_better\_random} and one \texttt{search\_better}, but not the other. The PCA plot further illustrate how the two algorithms finds the global maximum: \texttt{search\_better\_random} travels along the top right corner on the projected 2D circle while the path of \texttt{search\_better} is much shorter and only goes pass in between the two theoretical bases.

(there's more things going on about both algorithms experimenting with different seeds.)

@@

there's more text here Figure \ref{pca-better-random-tuned}, Comparison between search\_better\_random and two different search\_better. In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases. serach\_better\_random and one of the serach\_better have found the global maximum but the other fails. The lower panel shows the PCA plot for each of the three algorithms. The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum. The paths of two search\_better are relatively short and simple while the search\_better\_random has a more sophisticated path. This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance

@@

>>>>

Figure \ref{fig:toy-search} presents a comparison between the two plotting options discussed above. Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations. The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/toy-search-1.pdf}
\caption{\label{fig:toy-search}A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}. A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting. With \texttt{max\_tries\ =\ 500} specification, the algorithm is working much harder towards to the end.}
\end{figure}

@@

@@

@@

@@

The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is presented. The dataset used is \texttt{boa6} with \texttt{x2} and \texttt{x7} being informative. The two possible theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235 repectively and hence, the global maximum happens when variable \texttt{x7} is found. Simulation is done with 20 randomly generated seeds on different parameter specifications on two optimsers: \texttt{search\_better} and \texttt{search\_better\_random}. The data obbject is collected for each simulation and principal component analysis is applied to project the bases as points into 2D space with the orthonormality constraint imposed on the 6D space preserved by randomly generated basis on the surface of the 6D sphere.

Figure \ref{pca-better-default} shows that half of the seeds end up getting trapped in the local maximum on optimiser \texttt{search\_better}. This happens because \texttt{search\_better} will always accept a basis with higher index value. Figure \ref{pca-better-random-default} shows the results when the optimiser \texttt{search\_better\_random} is used. With a probability acceptance on inferior basis, more simulation ends up finds the global maximum. An evidence of the superior of this probabilistic implementation is simulation 5107 (row 1 column 2 in Figure \ref{pca-better-default} and row 3 column 5 in Figure \ref{pca-better-random-default}). When using \texttt{search\_better} to optimise, it goes straight down the track to find \texttt{V2} whilst with \texttt{search\_better\_random}, it first has a tendency to hand towards \texttt{V2} but then makes a sharp turn back to the track that finds \texttt{V7}.

In the next simulation, as in Figure \ref{pca-better-random-tuned}, the neighbourhood parameter \texttt{alpha} is increased from a 0.5 default to 0.7. Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated basis and hence represents a wider search. By this specification, all the 20 seends find the global maximum and a few simulations i.e.~seed 4761 (row 4, column 5), 9209 (row 3, column 2), 2986 (row 2, column 3) from this double guarantee of wider search and probabilistic acceptance.
